\documentclass[12pt]{letter}
\pagestyle{plain}
\sloppy

% Reduce space between \closing{ } and names
\makeatletter
\renewcommand{\closing}[1]{\par\nobreak\vspace{\parskip}%
  \stopbreaks
  \noindent
  \ifx\@empty\fromaddress\else
  \hspace*{\longindentation}\fi
  \parbox{\indentedwidth}{\raggedright
       \ignorespaces #1\\[2\medskipamount]%
       \ifx\@empty\fromsig
           \fromname
       \else \fromsig \fi\strut}%
   \par}
\makeatother
% 

\usepackage[headheight=200pt,headsep=200pt,left=1in,right=1in,top=0.8in,bottom=0.8in]{geometry}
\renewcommand{\baselinestretch}{1.0}
\addtolength{\headheight}{-10pt}
\addtolength{\textheight}{0pt}

\address{
Ben Southworth \\
Theoretical Division\\
Los Alamos National Laboratory\\
Los Alamos, NM 87544
\vspace{-5ex}
}
\date{}

\signature{Ben Southworth, Oliver Krzsik, and Will Pazner}

\begin{document}
\begin{letter}
\\
\opening{Dear Editor:}
\vspace{3ex}

We are writing to submit the manuscript
\begin{quote}
\emph{Fast parallel solution of fully implicit Runge-Kutta and discontinuous
  Galerkin in time for numerical PDEs, Part II: nonlinearities and DAEs},\\
by Ben Southworth, Oliver Krzsik, and Will Pazner
\end{quote}
for publication in the SIAM Journal on Scientific Computing (SISC).

This is the second paper of a pair that we are submitting to SISC simultaneously
focused on fast parallel solution of fully implicit Runge-Kutta (IRK) methods applied
to numerical PDEs. IRK methods have many benefits over
their more commonly used counterparts such as diagonally implicit Runge-Kutta (DIRK)
methods or BDF methods, but are rarely used in practice due to the difficulty
in solving the (non)linear stage equations. This paper develops new linearizations
as well as methods to solve the resulting linear systems using existing
solvers/preconditioners. Although some of the techniques used in Part I for
linear PDEs are not applicable here, we use a general theorem from Part I to
prove tight bounds on preconditioned condition numbers for simplified-Newton-type
linearizations, and further generalize this theory to provide an upper bound on
condition number for the linear preconditioning in more general nonlinear iterations.
We then demonstrate the new methods on more complex problems in compressible and
incompressible fluid dynamics.

Although IRK methods have seen significant research since the 1980s, few
papers have considered the efficient use of IRK methods for high-performance
simulation of numerical PDEs. We believe SISC is the appropriate journal for
this publication because it combines the development of both novel theory and
algorithms, making the use of IRK methods tractable, largely using existing
software infrastructure, and demonstrates the methods on interesting model
problems in fluid dynamics.

The article is original and has not been published in other forms. Please
contact us if further information is needed. We look forward to the
review.\\

\closing{Best Regards,}


\end{letter}
\end{document}
